Spatial considerations in the analysis of conservation data: evaluating the presence  of \textit{Acer campestre} (field maple) – space is special

### 2016-09-29 paper_v1.Rmd
Add files via upload

```{r eval=T, echo=F, message=F, warning=F, cache = TRUE}
# you will only need to run the lines below after 
# you have installed the packages on your computer
library(GISTools)
library(rgdal)
library(GWmodel)
library(repmis)

@@

Next, the presence of local collinearity amongst the predictor variables was tested. This is a critical step in any GWR analysis but one that is usually overlooked. Collinearity occurs when variables exhibit linear or near linear relationships. Strong collinearity will affect model reliability and precision, generate unstable parameter estimates, inflated standard errors and inferential biases (Dormann et al 2013), and there may be problems in separating variable effects (Meloun et al. 2002). In a GWR analyses, collinearity may occur locally, with the construction of localised regressions, even when it is not observed globally (Wheeler and Tiefelsdorf, 2005; Wheeler 2007, 2009, 2013; Brunsdon et al 2012). A number of approaches exist to address collinearity in regression modelling, such as partial least squares regression, principal component analysis regression and ridge regression (Hoerl 1962; Hoerl and Kennard 1970). Ridge regression is a penalised model where extensions, such as the lasso and the elastic net also provide predictor variable sub-set selection (e.g. Zou and Hastie 2005). All such models could be adapted to a localised form and Wheeler (2007; 2009) has proposed both ridge and lasso versions of GWR to address any detrimental local collinearity effects. A related, locally-compensated ridge GWR model is detailed in Brunsdon et al (2012), Lu et al. (2014b) and Gollini et al (2015). It has advantages over the ridge GWR model of Wheeler (2007) in that a ridge term is applied locally and not globally. These studies also describe associated local collinearity diagnostics for GWR, such as the use of local correlations amongst pairs of predictors, local Variance Inflation Factors (VIFs) for each predictor, local variance decomposition proportions (VDPs) and the local condition numbers (CNs). The key point about locally-compensated ridge GWR, is that a local ridge term is only applied where it is needed – when the local CN is above a pre-specified value in this case 30, which is a standard heuristic.

@@

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

```{r eval=T, echo=F, message=F, warning=F, cache = TRUE}
# you will only need to run the lines below after 
# you have installed the packages on your computer
library(GISTools)
library(rgdal)
library(GWmodel)
library(repmis)
library(knitr)

@@

Next, the presence of local collinearity amongst the predictor variables was tested. This is a critical step in any GWR analysis but one that is usually overlooked. Collinearity occurs when variables exhibit linear or near linear relationships. Strong collinearity will affect model reliability and precision, generate unstable parameter estimates, inflated standard errors and inferential biases (Dormann et al 2013), and there may be problems in separating variable effects (Meloun et al. 2002). In a GWR analyses, collinearity may occur locally, with the construction of localised regressions, even when it is not observed globally (Wheeler and Tiefelsdorf, 2005; Wheeler 2007, 2009, 2013; Brunsdon et al 2012). A number of approaches exist to address collinearity in regression modelling, such as partial least squares regression, principal component analysis regression and ridge regression (Hoerl 1962; Hoerl and Kennard 1970). Ridge regression is a penalised model where extensions, such as the lasso and the elastic net also provide predictor variable sub-set selection (e.g. Zou and Hastie 2005). All such models could be adapted to a localised form and Wheeler (2007; 2009) has proposed both ridge and lasso versions of GWR to address any detrimental local collinearity effects. A related, locally-compensated ridge GWR model is detailed in Brunsdon et al (2012), Lu et al. (2014) and Gollini et al (2015). It has advantages over the ridge GWR model of Wheeler (2007) in that a ridge term is applied locally and not globally. These studies also describe associated local collinearity diagnostics for GWR, such as the use of local correlations amongst pairs of predictors, local Variance Inflation Factors (VIFs) for each predictor, local variance decomposition proportions (VDPs) and the local condition numbers (CNs). The key point about locally-compensated ridge GWR, is that a local ridge term is only applied where it is needed – when the local CN is above a pre-specified value in this case 30, which is a standard heuristic.

@@

## 3.3 Code

All of the analyses and mappings were undertaken in R, the free open source statistical software. The Rmarkdown script used to produce this manuscript, including all the code used in the analysis and to produce the mapped figures, can be found at [https://github.com/lexcomber/SpatEcolPap](https://github.com/lexcomber/SpatEcolPap)

# 4. Results

## 4.1 Exploratory Regressions

A standard OLS regression models was undertaken and the resultant coefficient estimates and significance values are shown in Table 1 below. PET and Wilderness were found to be significant predictors of Acer campestre distributions at the 5% level (i.e. with a less than 95% chance of occurring randomly). Interestingly, in contrast to the findings of Coudun et al (2006), mean Autumn rainfall was not found to be significantly associated with the *Acer campestre* distributions.

```{r eval=T, echo=F, message=F, warning=F, cache = TRUE}
m <- lm(tree~pet+rain+wqi, data = data.frame(d.a))
#summary(m)
#coef(m)
tab <- summary(m)[4]
tab <- tab[[1]]
tab <- round(tab, 3)
rownames(tab) <- c("Intercept", "Mean annual PET", "Mean Autumn rainfall","Mean Wilderness Quality Index")
```

\begin{center}
{Table 1. The global regression co-efficient estimates.}
\end{center}

```{r eval=T, echo=F, message=F, warning=F}
library(knitr)
kable(tab, booktabs = T, format = 'markdown') 
```

```{r eval=T, cache = T, results = 'hide', echo=F, message=F, warning=F}
bw2 <- bw.gwr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, kernel = "bisquare", adaptive = TRUE, approach = "CV") 
# create GW regression points
hg <- spsample(os10m,3500,'hexagonal',offset=c(0.5,0.5)) # 7500 later?

# create gwr.mod 
m.g2 <- gwr.basic(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, regression.points = hg, bw = bw2, kernel = "bisquare", adaptive = TRUE)

# have a look at the model summary
#m.g2
```

```{r eval=T, echo=F, message=F, warning=F}
tab2 <- (m.g2$SDF@data[,1:4])
tab2 <- round(apply(tab2, 2, fivenum), 3)
rownames(tab2) <- c("Min", "1st Qu", "Median", "3rd Qu", "Max")
tab2 <- t(tab2)
rownames(tab2) <- c("Intercept", "Mean annual PET", "Mean Autumn rainfall","Mean Wilderness Quality Index")
```

\begin{center}
{Table 2. The variation of the coefficient estimates arising from a GWR analysis.}
\end{center}

```{r eval=T, echo=F, message=F, warning=F}
library(knitr)
kable(tab2, booktabs = T, format = 'markdown') 
```

The underlying theoretical framework provided by GWR tests for non-stationarity in processes and relationships between factors. A standard GWR analysis was undertaken and in this case an optimal adaptive bandwidth of `r bw2` data points was determined using a cross-validation procedure.

The local coefficient estimates from this GWR model are shown in Table 2. They indicate considerable variation around the median in the degree to which increases in the predictor variables are associated with Acer campestre distributions. For example, considering the inter-quartile ranges shows that, in some places:

- An increase in PET of 100 values is associated with a decrease of `r tab2[2,2] * 100` trees; 
- That each increase of 0.3 in the wilderness index is associated with a decrease of 1 tree (`r tab2[4,2]` * 0.3); 
But in other locations: 
- A decrease in PET of 100 values is associated with an increase of `r tab2[2,4] * 100` trees; 
-	That each increase of 0.5 in the wilderness index is associated with an increase of 1 tree (`r tab2[4,4]` * 0.5).

The local variation in coefficient estimates in Table 2 is in contrast to the global coefficient estimates in Table 1.

```{r eval=T, echo=F, message=F, warning=F}
# convert the gw.model SDF to a Spatial Points Data Frame
library(GISTools)
m.gr2 = SpatialPointsDataFrame(m.g2$SDF,data.frame(m.g2$SDF), proj4string = os.proj)
```

```{r, fig.width=9, echo = F, cache = T, eval=T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the mean annual PET coefficient estimates, with the context of the original PET and species data.")}
# PET and GW coefs for PET

par(mar = c(1,1,1,0))
par(mfrow = c(1,3)) 
cx <- 0.4
sh <-auto.shading(d.a$pet[!is.na(d.a$pet)], n = 6,
	cols=brewer.pal(7,'YlOrRd')[2:7])
plot(os10m, lwd = 0.5)
choropleth(d.a, v = d.a$pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topleft", sh = sh, box.col = NA, cex = 0.7)
title("Mean annual PET")

plot(os10m, lwd = 0.5)
plot(data.sp[os10m,], add = T, cex = 0.4, pch = 1, col = "#2525254C")
title("Data Points")

sh = auto.shading(m.gr2$d.a.pet, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.pet, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("PET Global Coeff: %s",round(coef(m)[2], 3))
title(tit)
plot(os10m, lwd = 0.5, add = T)
```

```{r, fig.width=9, echo = F, cache = T, eval=T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the Wilderness Quality Index  coefficient estimates, with the context of the original WQI and species data.")}

# WQI and GW coefs for WQI
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh <-auto.shading(d.a$wqi[!is.na(d.a$wqi)], n = 6,
	cols=brewer.pal(7,'YlOrRd')[2:7])
plot(os10m, lwd = 0.5)
choropleth(d.a, v = d.a$wqi, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topleft", sh = sh, box.col = NA, cex = 0.7)
title("Wildness Quality Index")

plot(os10m, lwd = 0.5)
plot(data.sp[os10m,], add = T, cex = 0.4, pch = 1, col = "#2525254C")
title("Data Points")

sh = auto.shading(m.gr2$d.a.wqi, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW WQI: ", B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("WQI Global Coeff: %s",round(coef(m)[4], 3))
title(tit)
plot(os10m, lwd = 0.5, add = T)
```

## 4.2 GWR Local Collinearity Diagnostics

The potential for detrimental effects due to local collinearity has been ignored by nearly all of the GWR analyses reported in the literature, regardless of domain or subject. Collinearity occurs when one predictor variable has a strong positive or negative relationship with another, typically when it is less than -0.8 or greater than +0.8. Critically, collinearity may be absent when calculated globally (ie from all the data values), but may be present locally when a subset of the data is considered, as is the case with a GWR analysis. Table 3 shows the results of applying GWR collinearity diagnostics to the GWR model above. This generates local VDPs and local CNs at the same scale (i.e. using the same adaptive bandwidth of 21 data points). The results are summarised in Table 3 and mapped in Figure 5. These indicate a high degree of local collinearity, with CNs greater than 30 and VDPs greater than 0.5 using standard heuristics (Belsley et al 1980; O’Brien 2007), locally. These values suggest that the application of a locally-compensated ridge GWR is warranted.

```{r eval=T, echo=F, message=F, warning=F}
library(GWmodel)
```

```{r eval=T, cache = T, echo=F, message=F, warning=F}
# Examine LOCAL Collinearity
gw.col <- gwr.collin.diagno(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, bw = bw2, kernel = "bisquare", adaptive = TRUE)
```

```{r eval=T, echo=F, message=F, warning=F}
### VIFs greater than 10 -> suggestive of collinearity
### CNs greater than 30 -> suggestive of collinearity
# summary(gw.col$SDF)
# VIFs & CNs for the variables considered
# summary(gw.col$SDF[,1:4])
tab3 <- gw.col$SDF@data[,1:4]
tab3 <- round(apply(tab3, 2, fivenum), 3)
rownames(tab3) <- c("Min", "1st Qu", "Median", "3rd Qu", "Max")
tab3 <- t(tab3)
rownames(tab3) <- c("Mean annual PET", "Mean Autumn rainfall","Mean Wilderness Quality Index", "Local CN")
```

\begin{center}
{Table 3. GWR collinearity measures, describing local CN and local VDPs for each predictor variable.}
\end{center}

```{r eval=T, echo=F, message=F, warning=F}
library(knitr)
kable(tab3, booktabs = T, format = 'markdown') 
```

\newpage

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The spatial distribution of the local CNs and the Mean annual PET and Mean Wilderness Quality VIFs.")}

par(mar = c(0,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.5

sh = auto.shading(gw.col$SDF@data[,4], n = 7, 
	cols = brewer.pal(7, "Reds"))
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,4], sh, pch = 19, cex = cx, add = T)
title("Local CN: all > 30!")
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.0f", cex = 0.7, title = "CN")
plot(os10m, lwd = 0.5, add = T)

sh = auto.shading(gw.col$SDF@data[,1], n = 5, 
	cols = brewer.pal(5, "Blues"))
sh$breaks <- c(2,5,7,10)
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,1], sh, pch = 19, cex = cx, add = T)
title("Local VIF (PET): some > 10")
plot(os10m, lwd = 0.5, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.1f", cex = 0.7, title = "VIF")

sh = auto.shading(gw.col$SDF@data[,3], n = 5, 
	cols = brewer.pal(5, "Greens"))
sh$breaks <- c(2,5,7,10)
plot(os10m, lwd = 0.5, add = F)
choropleth(gw.col$SDF, v = gw.col$SDF@data[,3], sh, pch = 19, cex = cx, add = T)
title("Local VIF (WQI): some > 10")
plot(os10m, lwd = 0.5, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.1f", cex = 0.7, title = "VIF")
```

## 4.3 Final GW analysis

```{r eval=T, cache = T, echo=F, results = 'hide', message=F, warning=F}
# Compensate for local collinearity
# using GW locally comnsated ridge regression function
# First determine the new bandwidth
# This tries to find the bandwidth to fit a ridge term in every location where the CN was greater than 30 
lcr.bw <- bw.gwr.lcr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, kernel = "bisquare", adaptive = TRUE, lambda.adjust = TRUE, cn.thresh = 30)
```

```{r eval=T, cache = T, echo=F, message=F, warning=F}
# Then run the locally compensated GWR 
lcr.m <- gwr.lcr(d.a$tree~d.a$pet+d.a$rain+d.a$wqi, data = d.a, regression.points = hg, bw = lcr.bw, kernel = "bisquare", adaptive = TRUE, lambda.adjust = TRUE, cn.thresh = 30)
#summary(lcr.m$SDF$Local_CN)
#summary(lcr.m$SDF)
# extract the SPDF for analysis
m.gr3 = SpatialPointsDataFrame(lcr.m$SDF,data.frame(lcr.m$SDF), proj4string = os.proj)
```

Having tested for and identified local collinearity, a locally compensated ridge GWR (LCR-GWR) was specified. This applies a GW regression but with a locally-compensated ridge term and fits local ridge regressions with their own ridge parameters (i.e., the ridge parameter varies across space), but only does this at locations where the local Condition Number is above a user-specified threshold. In this case the CN threshold was specified as 30. An optimal adaptive bandwidth of `r lcr.bw` data points was again determined using a cross-validation procedure. Figures 6 and 7 shows the spatial distribution of the original GWR coefficients, those determined using a locally compensated ridge GWR and a map of the differences between the two, for PET and for Wilderness Quality Index. In both cases there are large and potentially important differences between the coefficient estimates from the GWR and those from the LCR-GWR.

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The coefficient estimates of the degree to which mean annual PET predicts *Acer campestre* arising from the original GWR, a locally compensated ridge GWR and a map of GWR minus LCR-GWR coefficients.")}

# Map Original GWR coefficient estimates and LCR GWR ones
# PET and GW coefs for PET 
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh = auto.shading(m.gr2$d.a.pet, n = 7, 
	cols = brewer.pal(7, "Blues"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "GWR Coeff")
tit <- sprintf("GWR PET Coeffs")
title(tit)

plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr3, m.gr3$d.a.pet, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "LCR-GWR Coeff")
tit <- sprintf("LCR GWR PET Coeffs")
title(tit)

val <- (m.gr2$d.a.pet - m.gr3$d.a.pet)
sh = auto.shading(val, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, val, sh, pch = 19, cex = cx, add = T)
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = "Difference")
tit <- sprintf("LCR GWR vs GWR PET Coeffs")
title(tit)
```

```{r, fig.width=9, echo = F, eval=T, cache = T,  message=F, warning=F, fig.cap=fig$cap("", "The coefficient estimates of the degree to which Wilderness Quality Index predicts *Acer campestre* arising from the original GWR, a locally compensated ridge GWR and a map of GWR minus LCR-GWR coefficients.")}

# WQI and GW coefs for WQI
par(mar = c(1,1,1,1))
par(mfrow = c(1,3)) 
cx <- 0.4
sh = auto.shading(m.gr2$d.a.wqi, n = 7, 
	cols = brewer.pal(7, "Greens"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, m.gr2$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("GWR WQI Coeffs")
title(tit)

plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr3, m.gr3$d.a.wqi, sh, pch = 19, cex = cx, add = T)
txt=expression(paste("LCR GW Coeff: ",B[PET]))
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("LCR GWR WQI Coeffs")
title(tit)

val <- (m.gr2$d.a.wqi - m.gr3$d.a.wqi)
sh = auto.shading(val, n = 7, 
	cols = brewer.pal(7, "Spectral"))
plot(os10m, lwd = 0.5, add = F)
choropleth(m.gr2, val, sh, pch = 19, cex = cx, add = T)
txt= "Diff"
choro.legend(px = "topright", sh = sh, box.col = NA, fmt = "%2.2f", cex = 0.7, title = txt)
tit <- sprintf("LCR GWR vs GWR WQI Coeffs")
title(tit)
```

# 5. Discussion

In this paper a series of analyses were undertaken to demonstrate the application and value of explicitly spatial analyses, focusing on GWR. These develop local statistical models in order to test for spatial non-stationarity and are in contrast to standard, a-spatial, statistical approaches that assume the relationships between factors to be the same everywhere.

Additionally, the paper highlights the importance of considering and testing for local collinearity especially in spatial dependence models such as GWR, even where none is found to exist globally. In this analysis very strong evidence for local collinearity was found when the data were tested using local collinearity diagnostics. In most applications of GWR this critical step is missed. Where local collinearity is found a locally-compensated ridge GWR can be applied (Brunsdon et al. 2012). This only fits local ridge regressions at locations where the local CN is above a user-specified threshold. Gollini et al (2015) discuss alternative approaches for handling collinearity, but locally-compensated GWR models have the potential provide more accurate local coefficient estimates in the presence of collinearity than that found with a standard GWR model **REF needed?**.

The Geographically Weighted paradigm offers an attractive and coherent framework for many areas of applied geographical analysis. Geographically Weighted (GW) approaches develop local statistical models in order to test for spatial non-stationarity and are in contrast to standard, a-spatial, statistical approaches that assume the relationships between factors to be the same everywhere. They reflect what is commonly referred to as Tobler’s 1st Law of Geography (Everything is related to everything else, but near things are more related to each other) and an understanding of the world when it is viewed through ‘Geography Goggles’. These promote a vision in which the wearer is interested in how and where things vary, does not expect (statistical) relationships to be same everywhere, does not consider the world to be not normally distributed especially in space, but rather expects processes, relationships, processes, trends etc to vary spatially and to find clusters, hotspots, coldspots, etc.

These ideas are not new: quantitative geography in 1980s identified the need to move away from the whole map statistics, particularly Stan Openshaw’s group at Newcastle and Julian Besag’s at Durham but also Luc Anselin at Arizona. But it is important to re-state them now for a number of reasons. First, all data are spatial now (well perhaps not quite all!) but with advent of ubiquitous GPS, most records, datasets and data points have location attached to them. Second, location is not just another variable precisely because of the spatial non-stationarity observed in many processes, with the result that many phenomena are normally or randomly distributed, as predicated by classic statistical models. Third, the need to think spatially and to and to consider the spatial dimensions in a different way is given further salience by the increased access to and use of very powerful GIS software. This is increasingly resulting in instances of poor and inappropriate use of very powerful tools, but that is another story (see Comber et al., 2015). Finally, we simply observe that geography googles are not usually worn by researchers working in many areas of applied geography, especially conservation, environmental science and remote sensing, where the whole map statistic persists.

**More…?**

# 6. Conclusions
**TBC**

# Acknowledgements - **TBC**
Mark O’Connell Spatial Ecology and Conservation conference

# References - **TBC**

Akaike H (1973). Information Theory and an Extension of the Maximum Likelihood Princi-ple. In BN Petrov, F

Csaki (eds.), 2nd Symposium on Information Theory, pp. 267–281. Akademiai Kiado, Budapest.

Belsley DA, Kuh E, Welsch RE (1980). Regression Diagnostics: Identifying Influential Data and Sources of Collinearity. John Wiley & Sons.

Bowman A (1984). An Alternative Method of Cross-Validation for the Smoothing of Density Estimates. Biometrika, 71, 353–360.

Brunsdon C, Charlton M, Harris P (2012). Living with Collinearity in Local Regression Models. In Proceedings of the 10th International Symposium on Spatial Accuracy Assessment in Natural Resources and Environmental Sciences. Brasil.

Brunsdon C, Fotheringham AS, Charlton M (1996). Geographically Weighted Regression: A Method for Exploring Spatial Nonstationarity. Geographical Analysis, 28, 281–289.

Comber AJ, Brunsdon C, Green E. (2008). Using a GIS-based network analysis to determine urban greenspace accessibility for different ethnic and religious groups. Landscape and Urban Planning, 86: 103–114.

Comber A, Dickie J, Jarvis C, Phillips M and Tansey K, (2015). Locating bioenergy facilities using a modified GIS-based location-allocation-algorithm: considering the spatial distribution of resource supply. Applied Energy, 154: 309-316.

Coudun, C., Gégout, J.-C., Piedallu, C. and Rameau, J.-C. (2006), Soil nutritional factors improve models of plant species distribution: an illustration with Acer campestre (L.) in France. Journal of Biogeography, 33: 1750–1763

Dormann CF et al 2013. Collinearity: a review of methods to deal with it and a simulation study evaluating their performance. Ecography 36: 27-46.

Fischer & Gosset 1935

Fotheringham, A. S., Brunsdon, C., & Charlton, M. (2002). Geographically weighted regression: the analysis of spatially varying relationships. John Wiley & Sons.

Fujita, M. (1989). Urban Economic Theory, Land Use and City Size, Cambridge: Cambridge University Press.

Gandin 1965

Gollini, I., Lu, B., Charlton, M., Brunsdon, C., & Harris, P. (2015). GWmodel: an R package for exploring spatial heterogeneity using geogra- phically weighted models. Journal of Statistical Software, 63 (17), 1–50.

Hijmans, Robert J., and Jane Elith. Species distribution modeling with R. (2016). http://www.idg.pl/mirrors/CRAN/web/packages/dismo/vignettes/sdm.pdf

Hoerl AE (1962). Application of Ridge Analysis to Regression Problems. Chemical Engineering Progress, 58(3), 54–59.

Hoerl AE, Kennard RW (1970). Ridge Regression: Biased Estimation for Nonorthogonal Problems. Technometrics, 12(1), 55–67.

Kery M., B. Gardner, and C. Monnerat, 2010. Predicting species distributions from checklist data using site-occupancy models. Journal of Biogeography. 37: 1851–1862

Kolmogorov 1941

Krige 1951

Kuiters, A. T., van Eupen, M., Carver, S., Fisher, M., Kun, Z., & Vancura, V. (2011). Wilderness register and indicator for Europe. Final Report - http://ec.europa.eu/environment/nature/natura2000/wilderness/pdf/Wilderness_register_indicator.pdf

Lu, B., Harris, P., Charlton, M., & Brunsdon, C. (2014). The GWmodel R package: further topics for exploring spatial heterogeneity using geographically weighted models. Geo-spatial Information Science, 17(2), 85-101.

Matérn 1960

Matheron 1963

Meloun, M. et al. 2002. Crucial problems in regression modelling and their solutions. Analyst 127: 433–450.

O’Brien RM (2007). A Caution Regarding Rules of Thumb for Variance Inflation Factors. Quality & Quantity, 41(5), 673–690.

Phillips S.J. and J. Elith, 2011. Logistic methods for resource selection functions and presence-only species distribution models, AAAI (Association for the Advancement of Artificial Intelligence), San Francisco, USA.

Phillips, S.J., M. Dudik, J. Elith, C.H. Graham, A. Lehmann, J. Leathwick, and S. Ferrier. 2009. Sample selection bias and presence-only distri- bution models: implications for background and pseudo-absence data. Ecological Applications 19: 181-197.

Tanguy, M.; Dixon, H.; Prosdocimi, I.; Morris, D. G.; Keller, V. D. J. (2015). Gridded estimates of daily and monthly areal rainfall for the United Kingdom (1890-2014) [CEH-GEAR]. NERC Environmental Information Data Centre. http://doi.org/10.5285/f2856ee8-da6e-4b67-bedb-590520c77b3c

Trabucco, A., and Zomer, R.J. 2009. Global Aridity Index (Global-Aridity) and Global Potential Evapo-Transpiration (Global-PET) Geospatial Database. CGIAR Consortium for Spatial Information. Published online, available from the CGIAR-CSI GeoPortal at: http://www.csi.cgiar.org

VanDerWal J., L.P. Shoo, C. Graham and S.E. Williams, 2009. Selecting pseudo-absence data for presence-only distribution modeling: how far should you stray from what you know? Ecological Modelling 220: 589- 
594.

Ward G., T. Hastie, S.C. Barry, J. Elith and J.R. Leathwick, 2009. Presence- 
only data and the EM algorithm. Biometrics 65: 554-563.

Wheeler D (2007). Diagnostic Tools and a Remedial Method for Collinearity in Geographically Weighted Regression. Environment and Planning A, 39(10), 2464–2481.

Wheeler D (2009). Simultaneous Coefficient Penalization and Model Selection in Geographically Weighted Regression: the Geographically Weighted Lasso. Environment and Planning A, 41(3), 722–742.

Wheeler D (2013). Geographically Weighted Regression. In M Fischer, P Nijkamp (eds.), Handbook of Regional Science. Springer-Verlag.

Wheeler D, Tiefelsdorf M (2005). Multicollinearity and Correlation among Regression Co-efficients in Geographically Weighted Regression. Journal of Geographical Systems, 7(2), 161–187.

Yaglom 1955

Zou, H., & Hastie, T. (2005). Regularization and variable selection via the elastic net. Journal of the Royal Statistical Society: Series B (Statistical Methodology), 67(2), 301-320.
